Clustering time series based on probability distributions across temporal granularities

### 2021-10-30 main.Rmd
data design section done

\noindent Consider the scenario when all three granularities $g1$, $g2$, and $g3$ are responsible for distinguishing the designs. This implies that the patterns across each granularity will change significantly for at least one among the to-be-grouped designs. We consider different distributions across categories (as in Table \ref{tab:tab-dist-design} top) that will lead to different designs (as in Table \ref{tab:tab-dist-design} below). Figure \ref{fig:plot-3gran-new} shows the linear and cyclic representation of the simulated variable under these five designs. As could be seen from the plot, it is impossible to decipher the structural difference in the time series variable just by looking at the linear view. The difference in structure becomes quite clear when we see the distribution across cyclic granularities. Hence, for the consequent scenarios, only graphical displays across cyclic granularities are provided to emphasize the distributional differences in categories.

@@

\noindent This is the case where one granularity will remain the same across all designs. We consider the case where the distribution of $v$ would vary across levels of $g2$ for all designs, across levels of $g1$ for few designs and $g3$ does not change across designs. So $g3$ is not responsible for distinguishing across designs. Figure
\ref{#fig:gran2and1-clubbed}(left) shows the considered design.

@@

\noindent Here only one granularity is responsible for distinguishing the designs. Designs change significantly only for the granularity $g3$. Figure \ref{fig:gran2and1-clubbed}(right) shows this.

@@

The proposed methods could be extended when two granularities of interest interact and we are interested to group subjects based on the interaction of the two granularities. For example, consider a group having a different weekday, weekend behavior in summer months, but not across winter. This type of joint behavior across granularities wknd-wday and month-of-year can be discovered by examining the distribution across combination of categories for different interacting granularities. Hence, in this scenario, we consider combination of categories to be generated from different distributions. For simplicity, consider a case with just two interacting granularities $g1$ and $g2$ of interest. As opposed to the last case, where we could examine distributions across $l_{g_1} + l_{g_2} = 5$ individual categories, with interaction, we need to examine the distribution of $l_{g_1}*l_{g_2}=6$ combination of categories. Consider $4$ designs in Figure \ref{fig:interaction-gran} where different distributions are assumed for different combination of categories resulting in different designs. Design-1 has no change in distributions across $g1$ or $g2$, while Design-2 and Design-3 change across only $g1$ and $g2$ respectively. Design-4 changes across categories of both $g1$ and $g2$. Design-3 and Design-4 looks similar according to their relative difference between consecutive categories, but Design-4 also changes across facets, unlike Design-3 where all facets look the same.

@@

## Visual exploration

>>>>

\noindent Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction. Meaning that at least one of the designs to be categorized will have significantly different patterns for each granularity. In Table \ref{tab:tab-dist-design} (top), we explore various distributions across categories (as shown in Table Table \ref{tab:tab-dist-design} (bottom). Figure \ref{fig:plot-3gran-new} shows the simulated variable's linear and cyclic representations for each of these five designs. The structural difference in the time series variable is impossible to discern from the linear view. The shift in structure may be seen clearly in the distribution of cyclic granularities (Figure \ref{fig:plot-3gran-new} (right)). The following scenarios use solely graphical displays across cyclic granularities to highlight distributional differences in categories.

@@

\noindent This is the case where one granularity will remain the same across all designs. We consider the case where the distribution of $v$ would vary across levels of $g2$ for all designs, across levels of $g3$ for few designs and $g1$ does not change across designs. The proposed design is shown in Figure \ref{fig:gran2and1-clubbed}(left).

@@

\noindent Only one granularity is responsible for identifying the designs in this case. This is depicted in Figure \ref{fig:gran2and1-clubbed} (right) where only $g3$ affects the designs significantly.

@@

The proposed methods could be extended when two granularities of interest interact and we want to group subjects based on the interaction of the two granularities. Consider a group that has a different weekday and weekend behavior in the summer but not in the winter. This type of combined behaviour across granularities  can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (wknd-wday and month-of-year in this example). As a result, in this scenario, we analyse a combination of categories generated from different distributions. Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$. In contrast to the previous situation, when we could study distributions across $l_{g_1} + l_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $l_{g_1}*l_{g_2}=6$ combination of categories. Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs. Design-1 exhibits no change in distributions across $g1$ or $g2$, whereas Design-2 and Design-3 alter across only $g1$ and $g2$, respectively. Design-4 varies across both $g1$ and $g2$ categories. Design-3 and Design-4 appear similar based on their relative difference across consecutive categories, however Design-4 also changes across facets, unlike Design-3, which has all facets look the same.

@@

## Visual exploration of findings
